systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2006-02-16 paper/trunk/systemfit_other.tex
added a subsection about how to use "hausman.systemfit" in section "using systemfit"; corrected description of hausman test in section "other issues"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@600 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hausman test}

Hausman \citep{hausman78} developed a test for misspecification.
The null hypotheses of the test is that all exogenous variables are
uncorrelated with all disturbance terms.
Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
but only the 3SLS estimator is (asymptotically) efficient.
Under the alternative hypothesis the 2SLS estimator is consistent
but the 3SLS estimator is inconsistent.
The resulting test statistic, distributed as a
$\chi^2$, is the difference between the covariance of an
3SLS estimator and the covariance of an 2SLS estimator.
If the difference is (close to) zero, the model is correctly specified.
The Hausman test statistic is,
\begin{equation}
  m = \left( \hat{\beta}_2 - \hat{\beta}_3 \right)^{'}
      \left( Var \left[ \hat{\beta}_2 \right] -
             Var \left[ \hat{\beta}_3 \right] \right)
      \left( \hat{\beta}_2 - \hat{\beta}_3 \right)
\label{eq:hausman}
\end{equation}
where $\hat{\beta}_2$ and $Var \left[ \hat{\beta}_2 \right]$ are the estimated
coefficient and variance-covariance matrix from 2SLS estimation, and
$\hat{\beta}_3$ and $Var \left[ \hat{\beta}_3 \right]$ are the estimated
coefficients and variance covariance-matrix from 3SLS estimation.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hausman test}
\label{sec:hausman}

Hausman \citep{hausman78} developed a test for misspecification.
The null hypotheses of the test is that all exogenous variables are
uncorrelated with all disturbance terms.
Under this hypothesis both the 2SLS and the 3SLS estimator are consistent
but only the 3SLS estimator is (asymptotically) efficient.
Under the alternative hypothesis the 2SLS estimator is consistent
but the 3SLS estimator is inconsistent.
The Hausman test statistic is,
\begin{equation}
  m = \left( \hat{\beta}_2 - \hat{\beta}_3 \right)^{'}
      \left( Var \left[ \hat{\beta}_2 \right] -
             Var \left[ \hat{\beta}_3 \right] \right)
      \left( \hat{\beta}_2 - \hat{\beta}_3 \right)
\label{eq:hausman}
\end{equation}
where $\hat{\beta}_2$ and $Var \left[ \hat{\beta}_2 \right]$ are the estimated
coefficient and variance-covariance matrix from 2SLS estimation, and
$\hat{\beta}_3$ and $Var \left[ \hat{\beta}_3 \right]$ are the estimated
coefficients and variance covariance-matrix from 3SLS estimation.
Under the null hypotheses this test statistic has a
$\chi^2$ distribution with degrees of freedom equal to the number of
estimated parameters.
